---
title: 'Predicting Brexit Referendum Based On Polls: One Day Before Vote'
author: "Abhinav Dholepat"
date: "27/07/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Packages And Data 

```{r}
library(dslabs)
library(ggplot2) 
library(dplyr)
library(tidyverse)
```

```{r}
data("brexit_polls")
```

## Basic Graphs To Understand Data 

```{r}
brexit_polls %>% 
  ggplot(aes(x = enddate)) + 
  geom_line(aes(y = remain), color = "Blue") + 
  geom_line(aes(y = leave), color = "Red") +
  xlab("Time Poll Published") + 
  ylab("Blue is Remain, Red is Leave")
```

```{r}
brexit_polls %>% 
  ggplot(aes(enddate, spread)) + 
  geom_line(color = "Orange") + 
  xlab("Time Poll Published") + 
  ylab("Spread") 
```
```{r}
brexit_polls <- brexit_polls %>% 
  mutate(p = ((spread + 1) / 2)) %>% 
  mutate(se = sqrt((p * (1 - p)) / samplesize)) %>% 
  mutate(lower = (spread - (2 * se)), upper = (spread + (2 * se))) 

brexit_polls %>% 
  ggplot(aes(x = enddate)) + 
  geom_line(aes(y = lower), color = "Red") + 
  geom_line(aes(y = upper), color = "Blue") +
  xlab("Time Poll Published") + 
  ylab("95% Confidence Interval of Spread") 
```

## Wrangling And Adjusting Data 
The Code below does the following: 
1) It finds the true spread; meaning, it assumes that "undecided" voters will, in the end, have to choose "remain", "leave" or not vote at all, we assume that this occurs randomly. 
2) We use the true spread to construct a new confidence interval 
3) Finds crediable pollster (polling organizations who have conducted more than 7 polls) 
4) Filtering the poll data to one week before the vote, since people change their mind over time

```{r}

brexit_polls %>% 
  group_by(pollster) %>% 
  summarize(n())


data <- brexit_polls %>% 
  mutate(true_remain = (remain / (remain + leave)), true_leave = (leave / (remain + leave))) %>% 
  mutate(true_spread = (true_remain - true_leave)) %>% 
  mutate(true_p = ((true_spread + 1) / 2)) %>% 
  mutate(se = (2 * sqrt(((true_p) * (1 - true_p)) / samplesize))) %>% 
  mutate(true_lower = (true_spread - (1.96 * se)), true_upper = (true_spread + (1.96 * se))) 


final_data <- data %>% 
  group_by(pollster) %>% 
  filter(n() >= 7) %>%
  filter(enddate == max(enddate)) %>%
  ungroup() %>% 
  filter(enddate <= "2016-06-22")
```

```{r}
final_data %>% 
  ggplot(aes(enddate)) + 
  geom_line(aes(y = true_lower), color = "Blue") + 
  geom_line(aes(y = true_upper), color = "Red") + 
  xlab("Time Poll Pubished") + 
  ylab("95% Confidence Interval Of Spread") 
```

## Final Calculations 
Code below does the following: 
1) We are left with 11 polls. Since these polls are taken is different locations around the UK, we treat all 11 polls as 1 single poll. We now get a very large sample and new spread of opnion. 
2) We use this to calculate the 95% confidence interval of vote outcome 
3) We calculate probability of UK voting to leave the EU 
4) We calculate the t score by placing each of the 11 polls in a basket of polls and finding the probability of leaving the EU
```{r}
mean_spread <- final_data %>% 
  summarize(d = (sum(true_spread * samplesize)) / (sum(samplesize))) %>%  
  pull(d)

print("Mean Spread")
mean_spread

p_hat <- (mean_spread + 1) / 2 

N <- sum(final_data$samplesize) 

final_se <- 2 * sqrt((p_hat * (1 - p_hat)) / N)

print("Lower Bound Confidence Interval Of Vote Outcome")
mean_spread - (2 * final_se)
print("Upper Bound Confidence Interval Of Vote Outcome")
mean_spread + (2 * final_se)

### Probability of Voting to leave the EU 
print("Probability Of Leaving EU")
pnorm(0, mean_spread, final_se)

### t test, treating each poll as a part of a basket of polls ### 

avg <- mean(final_data$true_spread)
SSDD <- sd(final_data$true_spread)  
t <- (0 - avg) / (SSDD / sqrt(10)) 
print("Probability Of Leaving EU, t-score")
pt(t, 9) 
```